---
title: Extreme drought , host density , sex and bullfrogs influence fungal pathogens
  infection REDO
author: "Travis McDevitt-Galles"
date: "9/21/2017"
output: html_document
---

```{r setup, include=FALSE, echo=False}

library(ggmap)
library(devtools)
library(ggplot2)
library(tidyr)
library(rethinking)
library(lme4)
library(geosphere)
library(ggridges)

setwd("~/Desktop/Fall 2017/Lab meeting/Reading/Adams_etal._redone")

drought.df <- read.csv("Adams_et_al._Data.csv")

str(drought.df)
```


To start we will redo the mixed model for bd prevalence in the yellow legged frog, 
first we need to create a new column for bd presence

```{r}


# empty vector to store BD status
drought.df$BD.prez <- rep(NA, nrow(drought.df))

drought.df$BD.prez[ drought.df$Z.Spore.Equiv == 0 ] <- 0
drought.df$BD.prez[ drought.df$Z.Spore.Equiv > 0 ] <- 1 

```

Sweet now we have a new column that indicates whether or not the specimen was BD positive ( for now we are ignoring load we will not do that in bayesian)

first we need to subset for the species code RABO
```{r echo=T}

# sub setting for Rana boylii 
rabo.only <- subset( drought.df , drought.df$Species == "RABO")
    
# crearing a new variable for water year and survey event as a factor
 
rabo.only$fWater.year <- factor( rabo.only$Water.Year )

rabo.only$fSurvey.event <- factor ( rabo.only$Survey.Event)

## also need to drop larve due to lack of variaiton

rabo.only <- rabo.only[rabo.only$Sex.stage != "LARV", ]


final.model <- glmer( BD.prez ~ fWater.year + Sex.stage + Bullfrogs +
                         Bullfrog.Time + (1|fSurvey.event), data = rabo.only,
                     family = 'binomial' )

summary(final.model)

## killer we are off on the p-values for some reason but we are hitting the 
## estimated coefficients, lets now look at Bullfrog BD prevalence
## 


Raca.only <- subset(drought.df, drought.df$Species == "RACA")

Raca.only <- Raca.only[ Raca.only$Sex.stage != "F", ]

Raca.only$fSample.event <- as.factor(Raca.only$Survey.Event)

final.RACA <- glm( BD.prez ~ (Water.temp) + Sex.stage ,
                     family='binomial', data = Raca.only )

summary(final.RACA)
```


Sweet looks like we properly recreated to models lets recreate some of their 
graphs

```{r}

pos.bd <- subset( drought.df, drought.df$Z.Spore.Equiv > 0)

ggplot(pos.bd, aes( y = log10(Z.Spore.Equiv), x = Species )) +geom_boxplot() + theme_classic()


temp.range <- seq( 10,22, length.out = 100)

inter <- -25.6479 

temp.effect <- 1.8956    

pred <- rep(NA, 100)
    
for( i in 1:100){
    pred[i] <- inter + temp.range[i] * temp.effect
}

pred.trans <- boot::inv.logit(pred)

temp.plot <- ggplot( ) +  geom_line(aes( x= temp.range, y = pred.trans)) +
    theme_classic()

temp.plot <- temp.plot + geom_point(data= Raca.only,aes( x = Water.temp, y = BD.prez ,
                                            color = Sex.stage),size = 5 )

temp.plot + xlab("Water temperature") + ylab("BD Status")

```



Exploring how to do a guassian process model in Rethinking package


```{r, echo=T}

#### DEMO !!!!! 

Dmat

data(Kline2)
d <- Kline2
data(islandsDistMatrix)
d$society <- 1:10
mGP <- map2stan(
    alist(
        total_tools ~ dpois( mu ),
        log(mu) <- a + aj[society],
        a ~ dnorm(0,10),
        aj[society] ~ GPL2( Dmat , etasq , rhosq , 0.01 ),
        etasq ~ dcauchy(0,1),
        rhosq ~ dcauchy(0,1)
    ),
    data=list(
        total_tools=d$total_tools,
        society=d$society,
        Dmat=islandsDistMatrix),
    constraints=list(
        etasq="lower=0",
        rhosq="lower=0"
    ),
    warmup=1000 , iter=5000 , chains=4 )

tracerplot(mGP)



```


Ok cool, seems kind of straight forward , lets give it a shot


first lets run in our own data

```{r}

setwd("/Users/travismcdevitt-galles/Dropbox/Thesis/2nd Chapter/All_Years")

# Reading in data, both Infection and Site level data
odo_par <- read.csv("Odo_allYear_rm.csv")
odo_par <- odo_par[-1176,] # removing an NA

host_den <- read.csv("odo_per_net.csv")
site_sum <- read.csv("site_summary.csv")

# Combinding data sets
odonat_df <- plyr::join(odo_par,host_den, by=c("Site", "Year"))


# Subsetting rana only sites

site_lev <- read.csv("Site_vars.csv") 

odonat_df <- plyr::join ( odonat_df , site_lev, by=c("Site", "Year","Round"))

odo_rana1 <- subset(odonat_df, odonat_df$Rana == 1)

site.lat <- read.csv(file.choose())


## you should reorder these rows to match the larger data file site order and
##  provide each site with a unique id number
levels(site.lat$Site)

Euclid.site <- as.matrix(distm(site.lat[,2:3]))

colnames(Euclid.site) <- site.lat$Site
row.names(Euclid.site) <-  site.lat$Site

Euc <- Euclid.site/1000

## Ok i guess lets stop fuckind around and build the damn model


gau.pros.at1 <- map2stan(
    alist(
        Haem ~ dbinom(1, p), # haemtoloechus is distributed by probability "p" , the 1 is treating it like a bernoulli trial
        logit( p ) <- a + g[Site] , # just looking at how spaitials autocorrelated our data is
        g[Site] ~ GPL2( meterDist , etasq, rhosq, sigma ) , 
        a ~ dnorm(0, 10), # prio for our pooled mean risk
        etasq ~ dcauchy( 0 , 1), # prior for the max correlation
        rhosq ~ dcauchy( 0, 1), # prior for the decay rate
        sigma ~ dcauchy(0,1) # prior for the site level variance
    ),
    data = list(
        Haem = odo_rana1$Haem,
        Site = odo_rana1$Site,
        meterDist = Euc
    ),
    warmup = 2000, iter = 1e4, chains =4
)

summary(gau.pros.at1)


## huh it looks like it worked!!!! lets plot the autocorrelation


post <- extract.samples( gau.pros.at1 )

curve( (median(post$etasq))* exp(-median(post$rhosq)* x^2 ), from=0 ,
to=5,  ylim=c(0,10), ylab="Covariance", xlab="distance in Km", lwd=5) 


for(i in 1:100){
    curve((post$etasq[i])* exp(-(post$rhosq[i]) * x^2 ),add=T,
          col=col.alpha("black", .2) ) 
}



# extracting correlation matrix for the sites
# 
# 
    
K <- matrix(0 , nrow=32, ncol=32)

for( i in 1:32){
    for( j in 1:32){
        K[i,j] <- median(post$etasq) * 
            exp(-median(post$rhosq) * Euc[i,j]^2) 
    }
}

diag(K) <- median(post$etasq) + median(post$sigma)

Rho <- round( cov2cor(K), 2)

```


Ok so the hierarchaly guassian process didnt work, maybe you dove in
too deep, lets try something a bit easier and collapse the data to
the site level , ignoring everything else

```{r}

site_sum <- odo_rana1 %>%
    group_by(Site) %>%
    summarize( n = n(),
                prev.haem = mean(Haem) ,
               nHaem = sum(Haem),
               uninf = n() - sum(Haem),
                odo.den = mean(odo_per),
               pH = mean(pH)
    )



gau.pros.at2 <- map2stan(
    alist(
        Haem ~ dbinom(n, p), # haemtoloechus is distributed by probability "p" , the 1 is treating it like a bernoulli trial
        logit( p ) <- a + g[Site] , # just looking at how spaitials autocorrelated our data is
        g[Site] ~ GPL2( meterDist , etasq, rhosq, 0.01 ) , 
        a ~ dnorm(0, 10), # prio for our pooled mean risk
        etasq ~ dcauchy( 0 , 1), # prior for the max correlation
        rhosq ~ dcauchy( 0, 1) # prior for the decay rate
    ),
    data = list(
        Haem = site_sum$nHaem,
        Site = site_sum$Site,
        meterDist = Euc,
        n = site_sum$n
    ),
    warmup = 2000, iter = 1e4, chains =4
)


summary(gau.pros.at2)


post <- extract.samples( gau.pros.at2 )

curve( (median(post$etasq))* exp(-median(post$rhosq)* x^2 ), from=0 ,
to=5, ylim=c(0,10), ylab="Covariance", xlab="distance in Km") 

for(i in 1:100){
    curve((post$etasq[i])* exp(-(post$rhosq[i]) * x^2 ),add=T,
          col=col.alpha("black", .2) ) 
}


K <- matrix(0 , nrow=32, ncol=32)

for( i in 1:32){
    for( j in 1:32){
        K[i,j] <- median(post$etasq) * 
            exp(-median(post$rhosq) * Euc[i,j]) 
    }
}

diag(K) <- median(post$etasq) + 0.01

Rho <- round( cov2cor(K), 2)
View(Rho)


```

Cool lets look at additon factors in shaping this relationship


```{r}

GP_fish <- map2stan(
    alist(
        Haem ~ dbinom(1, p) ,
        logit( p ) <- a + g[Site] + t[genus]  + b1*Fish,  
        g[Site] ~ GPL2( meterDist , etasq, rhosq, sigma ) , 
        a ~ dnorm(0, 10), # prior for global intercept
        b1 ~ dnorm(0, 10), # prior fish effect
        t[genus] ~ dnorm( 0, 10), # prior for random effect of taxa
        etasq ~ dcauchy( 0 , 1), # prior for the max correlation
        rhosq ~ dcauchy( 0, 1), # prior for the decay rate
        sigma ~ dcauchy(0,1) # prior for the site level variance
    ),
    data = list(
        Haem = odo_rana1$Haem,
        Site = odo_rana1$Site,
        Fish = odo_rana1$Fish,
        meterDist = Euc,
        genus = odo_rana1$Genus
    ),
    warmup = 2000, iter = 1e4, chains =4
)



summary(GP_fish)

post1 <- extract.samples(GP_fish) 

gaus <- post1$g
class(gaus)
gaus <- as.data.frame(t(gaus))
gaus$Site <- site.lat$Site

Sites <- rep(site.lat$Site,32000)
Sites <- sort(Sites)

post.site <- as.vector(gaus)

joy.df <- cbind.data.frame(Sites, post.site)


ggplot(joy.df, aes( x=post.site, y= Sites )) + geom_density_ridges(scale=3) + 
    xlim(-5,5) + theme_ridges() + xlab("Gaussian adjustment")



gt <- post1$t

Genera <- rep(unique(odo_rana1$Genus), 32000 )
postit <- as.vector(gt)

post.tax <- cbind.data.frame(Genera, postit)

ggplot(post.tax, aes( x=postit, y= Genera )) + geom_density_ridges(scale=3) + 
   theme_ridges() + xlab("Gaussian adjustment") + xlim(-20,20)





# map of the sites
# 

bayMap <-qmap(location = "Sunol ,CA",zoom =9 )
bayMap + geom_point( aes( x= Longitude, y= Latitude), data= site.lat, col='red', size=1.5)
```